---
title: "First Deliverable"
author: "Javier Torralba"
date: "2025-02-07"
output:
  pdf_document: default
  html_document: default
---
# Telecom project
This project studies the merger of two of the four main telecommunication providers in the United States. The four main companies in this market are AT&T, Verizon, Sprint, and T-Mobile. In 2020, the merger of both Sprint and T-Mobile was approved, causing market changes. These market changes are still not known today. This project aims to fill the gap in the literature explaining the effects of the merger in the market shares of the American Telecommunication market. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load relevant libraries and set WD
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(tidyr)
library(viridis)
```

## Understanding the opening and closing of stores
We have data on the openings and closures of Sprint and T-Mobile stores in the US. These visualizations allow us to understand when each store opened and closed

```{r}
# Load data
ozgun_2018 <- fread("~/researchdrive/telecom/safegraph/ozgun/2018data.csv")
ozgun_2019 <- fread("~/researchdrive/telecom/safegraph/ozgun/2019data.csv") 
ozgun_2020 <- fread("~/researchdrive/telecom/safegraph/ozgun/2020data.csv")
base_file <- fread('~/researchdrive/telecom/safegraph/ozgun/base_file.csv')

# Loading relevan data
telecom_safegraph_2018 <- merge(ozgun_2018, base_file, by = 'placekey')
telecom_safegraph_2019 <- merge(ozgun_2019, base_file, by = 'placekey')
telecom_safegraph_2020 <- merge(ozgun_2020, base_file, by = 'placekey')

rm(ozgun_2018)
rm(ozgun_2019)
rm(ozgun_2020)

# Filtering for Sprint locations
sprint_2018 <- telecom_safegraph_2018 %>%
  filter(location_name == "Sprint")

sprint_2019 <- telecom_safegraph_2019 %>%
  filter(location_name == "Sprint")

sprint_2020 <- telecom_safegraph_2020 %>%
  filter(location_name == "Sprint")


# Relevant Sprint 2020 data
relevant_sprint_2020 <- sprint_2020 %>%
  select(placekey, 
         month, 
         year, 
         filename, 
         raw_visit_counts, 
         raw_visitor_counts,
         street_address,
         city,
         region,
         postal_code,
         opened_on, 
         closed_on,
         tracking_closed_since,
         postal_code) %>%
  mutate(count = 1) %>%
  group_by(placekey, closed_on, street_address, opened_on, postal_code) %>%
  summarise(sum(count)) %>%
  mutate(closed_on = ifelse(closed_on == "", "still open", closed_on))

```

```{r}
# Let's visualize the frequency of closed Sprint stores!
ggplot(relevant_sprint_2020, aes(x = closed_on)) + 
  geom_bar() +
  theme_bw() +
  labs(title = "Frequency of closing of Sprint stores",
       x = "Closed On",
       y = "Count",
       caption = "Safegraph data claims to have a 60 day margin of error. Made by Javier Torralba")

```

```{r}
ggplot(relevant_sprint_2020, aes(x = opened_on)) + 
  geom_bar() +
  theme_bw() +
  labs(title = "Frequency of opening of Sprint stores in data",
       x = "Opened On",
       y = "Count",
       caption = "Made by Javier Torralba")
```
```{r}
# Filtering for Sprint locations
t_mobile_2018 <- telecom_safegraph_2018 %>%
  filter(brands == "['T-Mobile']")

t_mobile_2019 <- telecom_safegraph_2019 %>%
  filter(brands == "['T-Mobile']")

t_mobile_2020 <- telecom_safegraph_2020 %>%
  filter(brands == "['T-Mobile']")


# Relevant Sprint 2020 data
relevant_t_mobile_2020 <- t_mobile_2020 %>%
  select(placekey, 
         month, 
         year, 
         filename, 
         raw_visit_counts, 
         raw_visitor_counts,
         street_address,
         city,
         region,
         postal_code,
         opened_on, 
         closed_on,
         tracking_closed_since) %>%
  mutate(count = 1) %>%
  group_by(placekey, closed_on, street_address, opened_on, region) %>%
  summarise(sum(count)) %>%
  mutate(closed_on = ifelse(closed_on == "", "still open", closed_on))
```

Visualize the data
```{r}
ggplot(relevant_t_mobile_2020, aes(x = opened_on)) + 
  geom_bar() +
  theme_bw() +
  labs(title = "Frequency of opening of T-Mobile stores",
       x = "Opened on",
       y = "Count",
       caption = "Safegraph data claims to have a 60 day margin of error. Made by Javier Torralba") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
ggplot(relevant_t_mobile_2020, aes(x = closed_on)) + 
  geom_bar() +
  theme_bw() +
  labs(title = "Frequency of closing of T-Mobile stores",
       x = "Closed on",
       y = "Count",
       caption = "Safegraph data claims to have a 60 day margin of error. Made by Javier Torralba") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Understanding market share
Given that we want to study the effects of the merger in the market, it is essential to understand the changes in market share from the point of descriptive statistics before and after the merger. 

```{r}
# Load Market Share data
test_my_df <- fread("~/researchdrive/telecom/Javiers_tests/test_merger_on_market_share_regression_analysis/fraction_of_market_share_per_cbsa/visualizations_market_share_across_time/mkt_share_data.csv")

# Create weighted average of market share data per CBSA (geo locator)
weighted_average_market_share <- test_my_df %>%
  group_by(date) %>%
  summarise(weighted_average_market_share_at_t = weighted.mean(market_share_per_cbsa_at_t, w = size_of_cbsa),
            weighted_average_market_share_verizon = weighted.mean(market_share_per_cbsa_verizon, w = size_of_cbsa),
            weighted_average_market_share_t_mobile = weighted.mean(market_share_per_cbsa_t_mobile, w = size_of_cbsa),
            weighted_average_market_share_sprint = weighted.mean(market_share_per_cbsa_sprint, w = size_of_cbsa)) %>%
  mutate(year = format(date, "%Y"),
         month = format(date, "%m")) %>%
  pivot_longer(cols = c(weighted_average_market_share_at_t, weighted_average_market_share_verizon, weighted_average_market_share_t_mobile, weighted_average_market_share_sprint),
               names_to = "companies",
               values_to = "market_share")

# Set month as factor
weighted_average_market_share$month <- factor(weighted_average_market_share$month, levels = sprintf("%02d", 1:12))
```

```{r}
# Visualize the market share over time!
ggplot(weighted_average_market_share, aes(x = month, y = market_share, group = year, color = year)) +
  geom_line(size = 0.7) +
  labs(x = "Month", 
       y = "Weighted average proportion of market share", 
       title = "Weighted average market share per company", 
       subtitle = "weight = size of CBSA",
       caption = "Made by Javier Torralba") +
  scale_color_viridis(discrete = TRUE, option = "inferno") +
  theme_bw() +
  facet_wrap(~companies, scales = "free_y")
```


